
# Executive Summary

Gulf of Alaska `r species` (*Sebastes polyspinis*) are classified as a Tier 3 stock and assessed using a statistical age-structured model with a full stock assessment typically produced in even years and a harvest projection stock assessment produced in odd years to coincide with the availability of new trawl survey data (odd years).
For Gulf of Alaska northern rockfish in `r year`, we present a operational update assessment and projection model to recommend harvest levels for the next two years.
This assessment consists of a population model, which uses survey and fishery data to generate a historical time series of population estimates, and a projection model, which uses results from the population model to predict future population estimates and recommended harvest levels. 
The data sets used in this assessment include total catch biomass, fishery age and size compositions, trawl survey abundance estimates, and trawl survey age compositions. 



## Summary of Changes in Assessment Inputs

*Changes in the input data:*  
Relative to the last full assessment the following substantive changes have been made to assessment inputs: 

  - Update survey biomass estimates through `r year-1`, 
  - Update survey age compositions with `r year-1` data,
  <!-- - Update fishery age compositions with `r year-2` data, -->
  - Update final catch values for `r year-2` and `r year-1`, and use preliminary catch for `r year`,
  - Update weight-at-age with survey data through `r year-1`,
  - Update size-at-age transition matrix with survey data through `r year-1`,
  - Update survey age composition input sample sizes to reflect current best practices.

*Changes in assessment methodology:*  

No model changes were implemented.

## Summary of Results

Model 24 was updated with new data ( `r pref_mod`). 
This model generally produces good visual fits to the data and biologically reasonable patterns of recruitment, abundance, and selectivity, and has a negative retrospective Mohn's rho value. 

*ABC recommendation*  
The `r pref_mod` projected age 2+ total biomass for `r year+1` is `r tot` t. 
The recommended ABC for `r year+1` is `r abc` t, the maximum allowable ABC under Tier `r tier`. 
This ABC is a `r perc`% increase compared to the `r year` ABC of `r abc0` and a `r perc1`% increase from the projected `r year+1` ABC from the prior assessment.
The `r year+1` GOA-wide OFL for `r species` is `r ofl` t.
The stock is not being subject to overfishing, is not currently overfished, nor is it approaching a condition of being overfished. 

\pagebreak

Reference values for GOA `r species` are summarized in the following table, with the recommended ABC and OFL values in bold.

```{r, child= "01a-main_table.Rmd", eval = T}

```


```{r, child= "01b-apportionment.Rmd", eval = T}

```




