---
params: 
  chapter: 10
  species: "northern rockfish"
  species_complex: "no"
  area: "Gulf of Alaska"
  tier: 3

author: "Benjamin C. Williams"
date: "November 2026"

# don't do anything below this line unless you are cool with the consequences
output:
  bookdown::html_document2:
     reference_docx: "template.docx"
     toc: false
     number_sections: false
     css: "styles.css"
     pandoc_args:  
       - '--lua-filter=scholarly-metadata.lua'
       - '--lua-filter=author-info-blocks.lua'

always_allow_html: true        
fontsize: 11pt
geometry:
  - top=1in
  - bottom=1in
  - left=1in
  - right=1in     
fig_caption: true
bibliography: "safe_bibby.bib"
csl: "fishfish.csl"

---

```{r setup, include=FALSE, echo = FALSE}
# remotes::install_github("crsh/rmdfiltr")
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.pos = "!h", out.extra = "")
```

```{r}
if(params$species_complex=="no"){
   title = paste0(params$chapter, ". Assessment of the ", tools::toTitleCase(params$species), " Stock in the ", params$area)
  } else {
    title = paste0(params$chapter, ". Assessment of the ", tools::toTitleCase(params$species), " Stock complex in the ", params$area)  
  }


```

---
title: `r title`
---

```{r load, warning=F, message=F}
# load
library(safe)
# globals
# utility function

pref_mod = folder = "m24-2026"
species = params$species
year = yr = 2026
source(here::here(year, 'r', 'utils.r'))
data_pull = read.table(here::here(year, "data", "raw", "data_called.txt"))[2,1]
data_pull2 = paste(months(as.Date(data_pull)), 
                    format(as.Date(data_pull,format="%Y-%m-%d"), format = "%d"))
catch = vroom::vroom(here::here(year, "data", "output", "fish_catch.csv"))
yld = vroom::vroom(here::here(year, "data", "output", "yld_rat.csv"))
prj = readLines(here::here(year, pref_mod, "proj", "model", "data", "nork_spcat.dat"))

c0 = comma(dplyr::filter(catch, year==yr-1)$catch)
c1 = comma(round(yld$proj_catch)) # estimated catch in current year
c2 = comma(stringr::str_sub(prj[grep(year+1, prj)], 6, 10)) # proj catch year + 1
c3 = comma(stringr::str_sub(prj[grep(year+2, prj)], 6, 10)) # proj catch yera + 2

exec_summ <- vroom::vroom(here::here(year, folder, "processed", "exec_summ.csv"))
exec_summ0 <- vroom::vroom(here::here(year-2, "m24", "processed", "exec_summ.csv"))

mod = readRDS(here::here(year, "m24_2026", 'm24_2026.rds'))
bf_dat = data.frame(age = 2:51,
                naa = mod$rpt$Nat[,66],
                waa = mod$rpt$waa,
                saa = mod$rpt$slx[,1])

best_f = afscassess::best_f(data = bf_dat, m=mod$rpt$M, last_ofl = 5848, last_f = 0.062) %>% 
  gsub(".*=", "", .) %>% 
  as.numeric(.) %>% 
  round(., 3)
tier = exec_summ[2,4]
tot = comma(exec_summ[3,4])

ofl = comma(exec_summ[11,4])
ofl0 = comma(exec_summ0[11,2])
sb1 = comma(exec_summ[4,4])
sb2 = comma(exec_summ[4,5])
b35 = comma(exec_summ[7,4])
b40 = comma(exec_summ[6,4])
b100 = comma(exec_summ[5,4])

f40 = exec_summ[9,4]
f35 = exec_summ[8,4]

# bring in as numeric
abc = as.numeric(exec_summ[13,4])
abc0 = as.numeric(exec_summ[13,2])
abc01 = as.numeric(exec_summ[13,3])

# calculate percents
perc = round((abc-abc0)/abc0 * 100)
perc1 = round((abc-abc01)/abc01 * 100)

# change to text

abc = comma(abc)
abc0 = comma(abc0)
```

This report may be cited as:

Williams, B.C. 2026. Assessment of the northern rockfish stock in the Gulf of Alaska. North Pacific Fishery Management Council, Anchorage, AK. Available from https://www.npfmc.org/library/safe-reports/

*This information is distributed solely for the purpose of pre-dissemination peer review under applicable information quality guidelines. It has not been formally disseminated by the National Marine Fisheries Service and should not be construed to represent any agency determination or policy.*

```{r, child="01-summary.Rmd", eval = T}
```

```{r, child="02-responses.Rmd", eval = T}
```

```{r, child="03-intro.Rmd", eval = F}
```

```{r, child="04-fishery.Rmd", eval = F}
```

```{r, child="05-data.Rmd", eval = T}
```

```{r, child="06-model.Rmd", eval = T}
```

```{r, child="07-results.Rmd", eval = T}
```

```{r, child="08-abc.Rmd", eval = T}
```

```{r, child="09-ecosystem.Rmd", eval = F}
```

```{r, child="10-gaps.Rmd", eval = F}
```
\pagebreak
\allsectionsfont{\centering}
# References
<div id="refs"></div>


\pagebreak

```{r, child="11-tables.Rmd", eval = F}
```
\pagebreak

```{r, child="12-figures.Rmd", eval = F}
```
\pagebreak

```{r, child="13-append_a.Rmd", eval = F}
```
\pagebreak

```{r, child="14-append_vast.Rmd", eval = F}
```
\pagebreak

```{r, child="15-append_compare.Rmd", eval = F}
```


