## Harvest recommendations


### Amendment 56 Reference Points

Amendment 56 to the GOA Groundfish Fishery Management Plan defines the “overfishing level”
(OFL), the fishing mortality rate used to set OFL ($F_{OFL}$), the maximum permissible ABC, and the fishing mortality rate used to set the maximum permissible ABC. 
The fishing mortality rate used to set ABC ($F_{ABC}$) may be less than this maximum permissible level, but not greater. 
Because reliable estimates of reference points related to maximum sustainable yield (MSY) are currently not available but reliable estimates of reference points related to spawning per recruit are available, Northern rockfish in the GOA are managed under Tier 3 of Amendment 56. 
Tier 3 uses the following reference points: $B_{40\%}$, equal to 40% of the equilibrium spawning biomass that would be obtained in the absence of fishing; $F_{35\%}$,,equal to the fishing mortality rate that reduces the equilibrium level of spawning per recruit to 35% of the level that would be obtained in the absence of fishing; and $F_{40\%}$, equal to the fishing mortality rate that reduces the equilibrium level of spawning per recruit to 40% of the level that would be obtained in the absence of fishing.
Estimation of the $B_{40\%}$ reference point requires an assumption regarding the equilibrium level of recruitment. 
In this assessment, it is assumed that the equilibrium level of recruitment is equal to the average of age-2 recruitments between 1979 and `r year-2`. 
Because of uncertainty in very recent recruitment estimates, we lag 2 years behind model estimates in our projection. 
Other useful biomass reference points which can be calculated using this assumption are $B_{100\%}$ and $B_{35\%}$, defined analogously to $B_{40\%}$. 
The `r year` estimates of these reference points are:

|$B_{100\%}$  |$B_{40\%}$   | $B_{35\%}$  | $F_{40\%}$ | $F_{35\%}$ 
|:----------|:--------------:|:-----------:|:---------:|:---------|
|`r b100`   |   `r b40`   | `r b35`       | `r f40`   | `r f35` |


### Specification of OFL and Maximum Permissible ABC

Female spawning biomass for `r year` is estimated at `r sb1` t. 
This is above the $B_{40\%}$ value of `r b40` t. 
Under Amendment 56, Tier 3, the maximum permissible fishing mortality for ABC is $F_{40\%}$ and fishing mortality for OFL is $F_{35\%}$. 
Applying these fishing mortality rates for `r year`, yields the following ABC and OFL:


|ABC | OFL  | 
|:----------|:--------------:|
|`r abc`   |   `r ofl`   | 

A standard set of projections is required for each stock managed under Tiers 1, 2, or 3 of Amendment 56.
This set of projections encompasses seven harvest scenarios designed to satisfy the requirements of Amendment 56, the National Environmental Policy Act, and the Magnuson-Stevens Fishery Conservation and Management Act (MSFCMA).

For each scenario, the projections begin with the vector of `r year` numbers at age as estimated in the assessment. 
This vector is then projected forward to the beginning of `r year + 1` using the schedules of natural mortality and selectivity described in the assessment and the best available estimate of total (year-end) catch for `r year`. 
In each subsequent year, the fishing mortality rate is prescribed on the basis of the spawning biomass in that year and the respective harvest scenario. 
In each year, recruitment is drawn from an inverse Gaussian distribution whose parameters consist of maximum likelihood estimates determined from recruitments estimated in the assessment. 
Spawning biomass is computed in each year based on the time of peak spawning and the maturity and weight schedules described in the assessment. 
Total catch after `r year` is assumed to equal the catch associated with the respective harvest scenario in all years. 
This projection scheme is run 1,000 times to obtain distributions of possible future stock sizes, fishing mortality rates, and catches.

Five of the seven standard scenarios will be used in an Environmental Assessment prepared in conjunction with the final SAFE. 
These five scenarios, which are designed to provide a range of harvest alternatives that are likely to bracket the final TAC for `r year+1`, are as follow (“$max F_{ABC}$” refers to the maximum permissible value of $F_{ABC}$ under Amendment 56):

 - Scenario 1: In all future years, *F* is set equal to $max F_{ABC}$. (Rationale: Historically, TAC has been constrained by ABC, so this scenario provides a likely upper limit on future TACs.)
 
 - Scenario 2: In `r year` and `r year + 1`, *F* is set equal to a constant fraction of $max F_{ABC}$, where this fraction is equal to the ratio of the realized catches in `r year-3`-`r year-1` to the ABC recommended in the assessment for each of those years. 
For the remainder of the future years, maximum permissible ABC is used. (Rationale: In many fisheries the ABC is routinely not fully utilized, so assuming an average ratio catch to ABC will yield more realistic projections.)

 - Scenario 3: In all future years, *F* is set equal to 50% of $max F_{ABC}$. (Rationale: This scenario provides a likely lower bound on $F_{ABC}$ that still allows future harvest rates to be adjusted downward when stocks fall below reference levels.)

 - Scenario 4: In all future years, *F* is set equal to the `r year-5`-`r year-1` average *F*. (Rationale: For some stocks, TAC can be well below ABC, and recent average *F* may provide a better indicator of $F_{TAC}$ than $F_{ABC}$.)

 - Scenario 5: In all future years, *F* is set equal to zero. (Rationale: In extreme cases, TAC may be set at a level close to zero.)

Two other scenarios are needed to satisfy the MSFCMA’s requirement to determine whether a stock is currently in an overfished condition or is approaching an overfished condition. 
These two scenarios are as follows (for Tier 3 stocks, the MSY level is defined as $B_{35\%}$):

 - Scenario 6: In all future years, *F* is set equal to $F_{OFL}$. (Rationale: This scenario determines whether a stock is overfished. If the stock is expected to be 1) above its MSY level in `r year` or 2) above ½ of its MSY level in `r year` and above its MSY level in `r year+10` under this scenario, then the stock is not overfished.)

 - Scenario 7: In `r year+1` and `r year+2`, *F* is set equal to max $F_{ABC}$, and in all subsequent years F is set equal to FOFL. (Rationale: This scenario determines whether a stock is approaching an overfished condition. If the stock is 1) above its MSY level in `r year+2` or 2) above 1/2 of its MSY level in `r year+2` and expected to be above its MSY level in `r year+12` under this scenario, then the stock is not approaching an overfished condition.)
 
Spawning biomass, fishing mortality, and yield are tabulated for the seven standard projection scenarios (Table `r params$chapter`-\@ref(tab:proj)). 
For projections in Scenario 2 (Author’s *F*); we use pre-specified catches to increase accuracy of short-term projections in fisheries where the catch is usually less than the ABC. 
This was suggested to help management with setting preliminary ABCs and OFLs for two-year ahead specifications.

In addition to the seven standard harvest scenarios, Amendments 48/48 to the BSAI and GOA Groundfish Fishery Management Plans require projections of the likely OFL two years into the future. 
While Scenario 6 gives the best estimate of OFL for `r year`, it does not provide the best estimate of OFL for `r year + 1`, because the mean `r year` catch under Scenario 6 is predicated on the `r year` catch being equal to the `r year` OFL, whereas the actual `r year` catch will likely be less than the `r year` OFL. 
The executive summary contains the appropriate one- and two-year ahead projections for both ABC and OFL.



## Risk Table and ABC recommendation


The SSC in its December 2018 minutes recommended that all assessment authors use the risk table when determining whether to recommend an ABC lower than the maximum permissible. 
The following template is used to complete the risk table:

```{r}
library(flextable)
library(tidyverse)

data.frame(level = c("Level 1: Normal",
                     "Level 2: Substantially increased concerns",
                     "Level 3: Major Concern",
                     "Level 4: Extreme concern"),
           assess = c("Typical to moderately increased uncertainty/minor unresolved issues in assessment.", 
                      "Substantially increased assessment uncertainty/ unresolved issues.",
                      "Major problems with the stock assessment; very poor fits to data; high level of uncertainty; strong retrospective bias.",
                      "Severe problems with the stock assessment; severe retrospective bias. Assessment considered unreliable."),
           pop = c("Stock trends are typical for the stock; recent recruitment is within normal range.",
                   "Stock trends are unusual; abundance increasing or decreasing faster than has been seen recently, or recruitment pattern is atypical.",
                   "Stock trends are highly unusual; very rapid changes in stock abundance, or highly atypical recruitment patterns.",
                   "Stock trends are unprecedented; More rapid changes in stock abundance than have ever been seen previously, or a very long stretch of poor recruitment compared to previous patterns."),
           enviro = c("No apparent environmental/ecosystem concerns",
                      "Some indicators showing adverse signals relevant to the stock but the pattern is not consistent across all indicators.",
                      "Multiple indicators showing consistent adverse signals a) across the same trophic level as the stock, and/or b) up or down trophic levels (i.e., predators and prey of the stock)",
                      "Extreme anomalies in multiple ecosystem indicators that are highly likely to impact the stock; Potential for cascading effects on other ecosystem components"),
           fish = c("No apparent fishery/resource-use performance and/or behavior concerns",
                    "Some indicators showing adverse signals but the pattern is not consistent across all indicators",
                    "Multiple indicators showing consistent adverse signals a) across different sectors, and/or b) different gear types",
                    "Extreme anomalies in multiple performance  indicators that are highly likely to impact the stock")) %>% 
  flextable() %>% 
    font(fontname = "Times", part = "all") %>% 
    width(j = ~ assess, width = 1.5) %>% 
    width(j = ~ pop, width = 2) %>% 
    width(j = ~ enviro, width = 1.5) %>% 
    width(j = ~ fish, width = 1.5) %>% 
    set_header_labels(level = "",
                    assess = "Assessment-related considerations",
                    pop = "Population dynamics considerations",
                    enviro = "Environmental/ecosystem considerations",
                    fish = "Fishery Performance") %>% 
  font(fontname = "Times", part = "all") %>% 
  fontsize(size = 10) %>% 
  theme_vanilla() %>% 
  flextable::fit_to_width(6.5) %>% 
  italic(part = "header") 
```  
The table is applied by evaluating the severity of four types of considerations that could be used to support a scientific recommendation to reduce the ABC from the maximum permissible. 
These considerations are stock assessment considerations, population dynamics considerations, environmental/ecosystem considerations, and fishery performance. Examples of the types of concerns that might be relevant include the following: 

  1. “Assessment considerations—data-inputs: biased ages, skipped surveys, lack of fishery-independent trend data; model fits: poor fits to fits to fishery or survey data, inability to simultaneously fit multiple data inputs; model performance: poor model convergence, multiple minima in the likelihood surface, parameters hitting bounds; estimation uncertainty: poorly-estimated but influential year classes; retrospective bias in biomass estimates.
  2. “Population dynamics considerations—decreasing biomass trend, poor recent recruitment, inability of the stock to rebuild, abrupt increase or decrease in stock abundance.
  3. “Environmental/ecosystem considerations—adverse trends in environmental/ecosystem indicators, ecosystem model results, decreases in ecosystem productivity, decreases in prey abundance or availability, increases or increases in predator abundance or productivity.
  4. “Fishery performance—fishery CPUE is showing a contrasting pattern from the stock biomass trend, unusual spatial pattern of fishing, changes in the percent of TAC taken, changes in the duration of fishery openings.”



#### Assessment considerations

Level 1. In recent assessments the GOA northern rockfish assessment model has resulted in a negative retrospective pattern, which is interpreted as the model continually increases spawning biomass as new data are added (-0.20 in 2018, -0.24 in 2020, and -0.082 in the current assessment, Figure 10-\@ref(fig:fretro)). 
While the assessment fits to composition data from the survey (age) and fishery (age) are generally adequate (Figures 10-\@ref(fig:ffac) and 10-\@ref(fig:fsac)), the fishery length compositions (Figure 10-\@ref(fig:ffsc)) are misaligned, though this fit has improved with the increase in the length plus group size. 
Changing from a design-based model to a VAST-based estimate has made the survey biomass estimates more realistic (less overall fluctuation) though the model continues to fit these data poorly. 
There is some question as to the efficacy of this trawl survey for developing indices of northern rockfish abundance. 
The items described here have been an issue for assessing northern rockfish for some time, we scored this category as Level 1, as the level of concern has not changed. 


#### Population dynamics considerations

Level 1. Recruitment since 2005 has been considerably lower than in 1970–2005. 
There is increasing proportions of GOA northern rockfish in the plus age groups for both survey and fishery age composition that indicates a substantial number of individuals are successfully surviving natural and fishing mortality to attain older ages and larger sizes. 
There is a reduction in body condition in recent years for young rockfish, though how this propagates through time is unclear. 
Skip spawning has been observed for this species, the spatial and temporal extent which is unknown.
However, preliminary investigations that incorporate skip spawning in maturity estimates lead to a reduction is spawning biomass and associated ABC.  
For these reasons we have given this risk table factor a level 2 concern for population dynamics considerations though make no recommendation for a reduction in ABC.


#### Environmental/Ecosystem considerations

Level 1. Environmental mechanisms for changes in survival remain unknown, though changes in water temperature and currents could have effects on prey abundance and success of transition of rockfish from pelagic to demersal stage. 
Predator effects would likely be more important on larval, post-larval, and small juvenile slope rockfish, but there is insufficient information on these life stages and their predators to inform a conclusion. 
Additionally, changes in bottom habitat due to natural or anthropogenic causes could alter survival rates by altering available shelter, prey, or other functions. 
Estimates of structural epifauna habitat (estimated using non-targeted data) have recently been in decline.
However, given the continued lack of biological and habitat information for northern rockfish, we scored this category as Level 1, as the level of concern has not changed.

#### Fishery performance

Level 1. Fishers usually direct their efforts first toward Pacific ocean perch because of its higher value relative to northern rockfish. 
After the TAC for Pacific ocean perch has been reached and NMFS closes directed fishing for this species, trawlers switch and target northern rockfish. 
The directed GOA northern rockfish fishery is concentrated on a limited number of highly productive locations. 
The patterns of fishing and percent of TAC taken have not substantially changed in the last three years, therefore we scored this category as Level 1.

#### Summary and ABC recommendation


```{r}
l1 = "Level 1: No increased concerns"
l2 = "Level 2: Substantially increased concerns"
l3 = "Level 3: Major concern"
l4 = "Level 4: Extreme concern"


data.frame(assess = l1,
           pop = l2,
           enviro = l1,
           fish = l1) %>% 
  flextable() %>% 
    width(j = ~ assess, width = 1.5) %>% 
    width(j = ~ pop, width = 1.5) %>% 
    width(j = ~ enviro, width = 1.5) %>% 
    width(j = ~ fish, width = 1.5) %>% 
    set_header_labels(level = "",
                    assess = "Assessment-related considerations",
                    pop = "Population dynamics considerations",
                    enviro = "Environmental/ecosystem considerations",
                    fish = "Fishery Performance") %>% 
  fontsize(size = 10) %>% 
  font(fontname = "Times", part = "all") %>% 
  italic(part = "header") 
```

We have ranked three categories as ‘Level 1: No apparent concern’ and one as a ‘Level 2, substantially increased concerns’. 
The GOA northerm rockfish assessment appears to fit available data well, the 2021 GOA trawl survey was undertaken as planned and data are included in this year’s assessment, and the fishery and environmental considerations appear to be within normal bounds. 
Because GOA northern rockfish ABC is has not been fully utilized in recent years we are not recommending a reduction in ABC at this time. 
We anticipate that we will monitor the survey abundance estimates, catch rates, and explore skip spawning more fully in the next assessment.

### Area Allocation of Harvests

Apportionment of ABC and OFL among regulatory areas has been based on the random effects model developed by the survey averaging working group. 
The random effects model was fit to the survey biomass estimates (with associated variance) for the Western, Central, and Eastern GOA. 
The random effects model estimates a process error parameter (constraining the variability of the modeled estimates among years) and random effects parameters in each year modeled. 
The fit of the random effects model to survey biomass in each area is shown in Figure 10-\@ref(fig:fre). 

In general the random effects model fits the area-specific design-based survey biomass estimates reasonably well. 
Based on the random effects estimates the area apportionments for GOA northern rockfish are 52.7% for the Western area (up from 37.76% in 2021), 47.3% for the Central area (down from 62.22% in 2018), and 0.02% for the Eastern area (same as 2021). 
The changes are due to the more frequent catches of northern rockfih in the western GOA during the 2021 survey Figure 10-\@ref(fig:fcpue).
Applying the random effect model apportionments to the recommended ABC for northern rockfish results in 2,614 t for the Western area, 2,350 t for the Central area, and 1 t for the Eastern area for `r year+1`.
For management purposes, the small ABC of northern rockfish in the Eastern area is combined with the Other Rockfish complex.


### Status Determination

Under the MSFCMA, the Secretary of Commerce is required to report on the status of each U.S. fishery with respect to overfishing. 
This report involves the answers to three questions: 1) Is the stock being subjected to overfishing? 2) Is the stock currently overfished? 3) Is the stock approaching an overfished condition?

*Is the stock being subjected to overfishing?* The official catch estimate for the most recent complete year (`r year-1`) is `r c0`  t. 
This is less than the `r year-1` OFL of `r ofl0` t. 
Therefore, the stock is not being subjected to overfishing.

Harvest Scenarios #6 and #7 are intended to permit determination of the status of a stock with respect to its minimum stock size threshold (MSST). 
Any stock that is below its MSST is defined to be overfished. 
Any stock that is expected to fall below its MSST in the next two years is defined to be approaching an overfished condition. 
Harvest Scenarios #6 and #7 are used in these determinations as follows:

*Is the stock currently overfished?* This depends on the stock’s estimated spawning biomass in `r year`:

a. If spawning biomass for `r year` is estimated to be below ½ $B_{35\%}$, the stock is below its MSST.  
b. If spawning biomass for `r year` is estimated to be above $B_{35\%}$ the stock is above its MSST.  
c. If spawning biomass for `r year` is estimated to be above ½ $B_{35\%}$ but below $B_{35\%}$, the stock’s status relative to MSST is determined by referring to harvest Scenario #6 (Table 10-\@ref(tab:proj)).

If the mean spawning biomass for `r year+12` is below $B_{35\%}$, the stock is below its MSST. 
Otherwise, the stock is above its MSST.

*Is the stock approaching an overfished condition?* This is determined by referring to harvest Scenario #7:

a. If the mean spawning biomass for `r year+2` is below 1/2 $B_{35\%}$, the stock is approaching an overfished condition.  
b. If the mean spawning biomass for `r year+2` is above $B_{35\%}$, the stock is not approaching an overfished condition.  
c. If the mean spawning biomass for `r year+2` is above 1/2 $B_{35\%}$ but below $B_{35\%}$, the determination depends on the mean spawning biomass for `r year+12` 
If the mean spawning biomass for `r year+12` is below $B_{35\%}$, the stock is approaching an overfished condition. 
Otherwise, the stock is not approaching an overfished condition.

Based on the above criteria and Table 10-\@ref(tab:proj), the stock is not overfished and is not approaching an overfished condition. 
The fishing mortality that would have produced a catch for last year equal to last year’s OFL is `r best_f`.
