# Executive Summary

Gulf of Alaska `r species` (*Sebastes polyspinis*) have historically been assessed on a biennial stock assessment schedule to coincide with the availability of new trawl survey data (odd years). 
In 2017, the Alaska Fisheries Science Center (AFSC) participated in a stock assessment prioritization process. 
It was recommended that the Gulf of Alaska (GOA) `r species` remain on a biennial stock assessment schedule with a full stock assessment produced in even years and a partial stock assessment produced in odd years. 
For a partial assessment, the projection model is updated with new catch information and results are used to recommend harvest levels for the next two years. 
This incorporates the most current catch information without re-estimating model parameters and biological reference points.

The GOA `r species` is classified as a Tier 3 stock and is assessed using a statistical age-structure model.
This assessment consists of a population model, which uses survey and fishery data to generate a historical time series of population estimates, and a projection model, which uses results from the population model to predict future population estimates and recommended harvest levels. 
The data used in this assessment includes total catch biomass, fishery age and size compositions, trawl survey abundance estimates, and trawl survey age compositions. 

## Summary of Changes in Assessment Inputs

*Changes in the input data:* There were no changes made to the assessment model inputs as this is an off-cycle year. 
New data added to the projection model included updated catch data from `r year-1` (`r filter(catch, Year == year-1) %>% pull(Catch) %>% prettyNum(., big.mark = ",", digits = 0)` t) and new estimated catches for `r year`-`r year + 2`. 
Catch data were queried on `r data_pull`. 
The `r year` catch was estimated by increasing the observed catch by an expansion factor of `r yld %>% pull(catch_rat) %>% round(., 3)`, which accounts for the average fraction of catch taken after `r data_pull2` in the last three complete years (`r year - 3`-`r year-1`). 
This expansion factor results in an estimated catch for `r year` of `r filter(catch, Year == year) %>% pull(Catch) %>% prettyNum(., big.mark = ",", digits = 0)` t. 
To estimate future catches, we updated the yield ratio to `r yld %>% pull(yld) %>% round(., 2)`, which was the average ratio of catch to ABC for the last three complete catch years (`r year - 3`-`r year-1`). 
This yield ratio was multiplied by the projected ABCs from the updated projection model to generate catches of **3,089** t in `r year + 1` and **2,884** t in `r year + 2`.

*Changes in assessment methodology:* There were no changes from the `r year-1` assessment [@Williams2022] since this is an off-cycle year.


## Summary of Results

*ABC recommendation*  
For the `r year + 1` fishery, we recommend maximum allowable ABC of **5,147** t from the updated projection model. 
This ABC is a **minor increase** over the `r year+1` projected ABC of **5,100** t from the 2022 full assessment. 
The 2022 GOA-wide OFL is 6,143 t.

The stock is not being subject to overfishing, is not currently overfished, nor is it approaching a condition of being overfished. 
The test for determining whether overfishing is occurring is based on the `r year -1` catch compared to OFL. 
The official total catch for `r year-1` is `r filter(catch, Year == year-1) %>% pull(Catch) %>% prettyNum(., big.mark = ",", digits = 0)` t which is less than the `r year -1` OFL of 6,396 t; therefore, the stock is not being subjected to overfishing. 
The tests for evaluating whether a stock is overfished or approaching a condition of being overfished require examining model projections of spawning biomass relative to $B_{35\%}$ for `r year + 1` and `r year + 2`. 
The estimates of spawning biomass for `r year + 1` and `r year + 2` from the current year projection model are** 40,474** t and **37,408** t, respectively. 
Both estimates are above the $B_{35\%}$ estimate of **29,691** t and, therefore, the stock is not currently overfished nor approaching an overfished condition.

Reference values for northern rockfish are summarized in the following table, with the recommended ABC and OFL values in bold.

```{r, child= "01a-main_table.Rmd", eval = T}

```


Text table of area apportionments (if any) for the recommended one- and two-year ahead ABCs and OFLs, with a brief description of the apportionment methodology.

```{r, child= "01b-apportionment.Rmd", eval = T}

```


If so are so kind as to provide tables for the plan team, place them here

```{r, child= "01c-plan_team.Rmd", eval = T}

```


